### 2024-04-17 paper/paper.qmd
renamed figures to match our paper and finish the rough draft for the data section

---
title: "Gamification Effectiveness on Education"
author: 
  - Sachin Chhikara
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
```

# Introduction

Games have been part of human history for a long time and are loved by many young and old. The design behind why game are so fun have investigated in recent time and how to apply those element that are fun and to apply them to other fields, this is known as gamification. Gamification is used in fields such as enterprise, sales, lifestyle and education. In this paper we will focusing on application of education, we be using data from Duolingu, training model and making inference off, real life application of gamification to an individual productive.

In this paper, we are doing replication of a study ... , which has a control and treatment where the treatment is applied gamification, while the control is using an ordinary education system. Yeah there benefits, but there are downside when comes to how the gamification as it might discourage learning. Furthermore, an wide use application of gamification is a language learning app called Duolingo, which has ... active users and this app learning is recognized in certain Universities as legitimate English learning.So we examine game design that are use. Ways to apply gamification in daily lives, like this app called Habitica: Gamify your tasks.

# Data

## Data Source

This paper reproduces the result from called "On the effectiveness of game-like and social approaches in learning: Comparing educational gaming, gamification & social networking" (@DEMARCOS2017155). This paper will provided a overview of the data used in that study. For more detailed explanation of the data please visit the original study.

## Methodology

The original study used 379 undergraduates for their experiments and there were randomly assigned into 5 groups Control(N=76), Educational game(N=75), Gamification Plugin(N=77), Social networking website(N=75), Social gamified networking website(N=76). The experiment was carried out on 10-week first year undergraduate class called ‘Qualification for ICT Users’ with hybrid learning. This course is an introduction to basic computing concepts such as word processing, spreadsheets, presentations and databases. There is an assignment for each of these four concepts which focus on testing practical skills. While, the final examination is cumulative and focus on testing conceptual knowledge.

Students have lecture every two weeks and must work independently rest of the work to learn then material. Supplementary learning material was provided such as videos and communication tools in learning material platform.

The control group just use what was mention above. Educational game used was called Ribbonhero offered by Microsoft as as free download to download Software, and player plays as virtual character Clippy and must overcome six challenges with increasing difficulty. Gamification plugin (@fig-Gamification-plugin) is integrated in the institutional management system where there are challenges, trophies/badges and a leader board. Social Networking website was run on a open source engine, Elgg (@fig-social_network) and it functions very similar to social media where students can share their thoughts, make friends and discuss the material in class. Social gamified networking website use the same engine as the previous group and having similar features to Gamification plugin(@fig-social_gamification). A summary of the groups features and potential benefits(@tbl-summary-group). The study highlights that the educational game is not necessarily aligned with the learning goals, while the gamification used in this study was.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Gamification-plugin
#| fig-cap: "Gamifcation Plugin. Leaderboard and challenges."
knitr::include_graphics(file.path(here::here(),"other/figures/figure1.jpg")) 
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-social_network
#| fig-cap: "Social Networking Website."
knitr::include_graphics(file.path(here::here(),"other/figures/figure2.jpg")) 
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-social_gamification
#| fig-cap: "Social gamification website: Gamification tools (left), dashboard (top-right) & leaderboard (bottom-right)."
knitr::include_graphics(file.path(here::here(),"other/figures/figure3.jpg")) 
```
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-summary-group
#| tbl-cap: "Summary of the groups used, their features and potential benefits."

htmltools::includeHTML(file.path(here::here(),"other/tables/table1.html"))
```

## Data Cleaning Procedure

For our replication, we drop 20 rows of data from the original dataset provided since those rows contained missing data on grades of some assessments and final examination. The age row was removed since in the original study it mention that there age range for each group but in the dataset provided they made every value in that column to 19 since that the average age in the study. To summarize, the size of the dataset we use in this replication is N = 359(Control=75, Game=66, Plugin=73, Social Network=74, Social Gamification=71).

This dataset was chosen since my many studies in gamification don't have easy access to their datasets. In addition, this dataset includes control and treatment group which is useful for examining the difference between a traditional education system and gamification.

## Data Visualization

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-pretest
#| tbl-cap: "The results of the pretest."

htmltools::includeHTML(file.path(here::here(),"other/tables/table2.html"))
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-posttest
#| tbl-cap: "The results of the posttest."

htmltools::includeHTML(file.path(here::here(),"other/tables/table3.html"))
```
The variables that are we are interested in seeing if share a relationship is posttest of Word Processing, Spreadsheets, Presentations, Databases and final examination. And we are interested in seeing the difference between the pretest(@tbl-pretest) and posttest(@tbl-posttest) and how each group is affect by the experiment.

The results from pretest(@tbl-pretest) are pretty such the same for each group which aligns with the fact that this study was given before the experiment began. When looking at the post test results(@fig-boxplot), the control group is lagging behind the other groups, especially in word processing, spreadsheets and databases.

In word processing post test(@fig-boxplot), the group that perform the best were the social network and social gamifcation and roughly having the same performance. And then followed by educational game and gamification plugin. So it seems for word processing that groups with social elements are performing better than the group with game elements.

In spreadsheets post test(@fig-boxplot), the group that perform the best was the social gamification, while the other three groups perform about the same and was only little bit behind social gamification.

In presentations post test(@fig-boxplot), the group that perform the best was the educational game and not far behind was social network and gamification plugin. Surprisingly, social gamification was behind those three groups, as it would seem that the combination of social and game elements would make better than social network and gamification plugin.

In database post test(@fig-boxplot), the educational game performed as worse the control group. And similarly to the presentation posttest the social gamification group does worse than the social network and gamification plugin. And gamification plugin does slightly better than social network.

For the final examination(@fig-final-examination), interestingly enough, the group with the best results was the control group. Followed closely, by the social network group and then Educational game group. So the groups that performed the worse was gamfication plugin group and social gamfication group.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-boxplot
#| fig-cap: "Boxplot showing the results of the post test for control and experiment groups "
knitr::include_graphics(file.path(here::here(),"other/figures/figure4.png"))

```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-final-examination
#| fig-cap: "Boxplot showing the results of the final examination for each group that particapte in this experiment"
knitr::include_graphics(file.path(here::here(),"other/figures/figure5.png"))

```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

---
nocite: '@*'
---