### 2024-04-19 paper/paper.qmd
finish the dicussion section

---
title: "Gamification Effectiveness on Education"
author: 
  - Sachin Chhikara
thanks: "Code and data are available at: https://github.com/SachinChhikara/Effectiveness_of_Gamification_Education."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
```

# Introduction

Games have been part of human history for a long time and are loved by many young and old. The design behind why game are so fun have investigated in recent time and how to apply those element that are fun and to apply them to other fields, this is known as gamification. Gamification is used in fields such as enterprise, sales, lifestyle and education. In this paper we will focusing on application of education, we be using data from Duolingu, training model and making inference off, real life application of gamification to an individual productive.

In this paper, we are doing replication of a study ... , which has a control and treatment where the treatment is applied gamification, while the control is using an ordinary education system. Yeah there benefits, but there are downside when comes to how the gamification as it might discourage learning. Furthermore, an wide use application of gamification is a language learning app called Duolingo, which has ... active users and this app learning is recognized in certain Universities as legitimate English learning.So we examine game design that are use. Ways to apply gamification in daily lives, like this app called Habitica: Gamify your tasks.

# Data
## Data Source
This paper reproduces the result from called "On the effectiveness of game-like and social approaches in learning: Comparing educational gaming, gamification & social networking" (@DEMARCOS2017155). This paper will provided a overview of the data used in that study. For more detailed explanation of the data please visit the original study.

## Methodology
The original study used 379 undergraduates for their experiments and there were randomly assigned into 5 groups Control(N=76), Educational game(N=75), Gamification Plugin(N=77), Social networking website(N=75), Social gamified networking website(N=76). The experiment was carried out on 10-week first year undergraduate class called ‘Qualification for ICT Users’ with blended learning. This course is an introduction to basic computing concepts such as word processing, spreadsheets, presentations and databases.

In this experiment, before applying the corresponding learning methods, the students take a pretest, which occurs in week 1. And through week 2-9, depending on the group the student was assigned they learn using that specific method. During those weeks, they were assigned four assignments(aka post tests) on week 3, 5, 7 and 9. In addition, these assignments are on word processing, spreadsheets, presentations and databases, which tests students on practical skills. Finally, there is cumulative final exam and tests students on conceptual knowledge.

All groups had students have lecture every two weeks and must work independently rest of the work to learn then material using their respective method. In addition, Supplementary learning material was provided such as videos and communication tools in learning material platform. The control group just use traditional e-learning approach that was mention above. Educational game used was called Ribbonhero offered by Microsoft as as free download to download Software, and player plays as virtual character Clippy and must overcome six challenges with increasing difficulty. Gamification plugin (@fig-Gamification-plugin) is integrated in the institutional management system where there are challenges, trophies/badges and a leader board. Social Networking website was run on a open source engine, Elgg (@fig-social_network) and it functions very similar to social media where students can share their thoughts, make friends and discuss the material in class. Social gamified networking website use the same engine as the previous group and having similar features to Gamification plugin(@fig-social_gamification). A summary of the groups features and potential benefits(@tbl-summary-group). The study highlights that the educational game is not necessarily aligned with the learning goals, while the gamification used in this study was.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Gamification-plugin
#| fig-cap: "Gamifcation Plugin. Leaderboard and challenges."
knitr::include_graphics(file.path(here::here(),"other/figures/figure1.jpg")) 
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-social_network
#| fig-cap: "Social Networking Website."
knitr::include_graphics(file.path(here::here(),"other/figures/figure2.jpg")) 
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-social_gamification
#| fig-cap: "Social gamification website: Gamification tools (left), dashboard (top-right) & leaderboard (bottom-right)."
knitr::include_graphics(file.path(here::here(),"other/figures/figure3.jpg")) 
```
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-summary-group
#| tbl-cap: "Summary of the groups used, their features and potential benefits."

knitr::include_graphics(file.path(here::here(),"other/tables/table1.jpg"))
```

## Variables of Interest and Data Cleaning Procedure

The variables used in this study are Group, Gender, pretest and post test for Word Processing, Spreadsheets, Presentation, Databases and FinalExamination. Pretest, post test and FinalExamination use the measurement of a grade, which is from 0 - 100%.

For our replication, we drop 20 rows of data from the original dataset provided since those rows contained missing data on grades of some assessments and final examination. The age row was removed since in the original study it mention that there age range for each group but in the dataset provided they made every value in that column to 19 since that the average age in the study. To summarize, the size of the dataset we use in this replication is N = 359(Control=75, Game=66, Plugin=73, Social Network=74, Social Gamification=71).

This dataset was chosen since my many studies in gamification don't have easy access to their datasets. In addition, this dataset includes control and treatment group which is useful for examining the difference between a traditional education system and gamification.

# Model

Here we briefly describe the Bayesian analysis model used to investigate final examination marks and how it is influence by the group chosen for the student and the gender of the student. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the the final examination mark, which is a number 0 and 100. $beta_0$ is the intercept and it is the expected value of the control group. $group_i$ is defined as the experiment groups: educational game, gamification plugin, social network and social gamification. $gender_i$ is defined is binary value that is either male or female.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{Group}_i \times \mbox{Gender}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We used normal distribution for this model since our sample size was big enough(over 300 sample points), this allows to use the central limit theorem. We pick variable Group because we want to see how each experiment group hold up against our control group and we want to see if the game elements and social elements improve academic performance. We pick the gender variable since we want to see if there is a learning method that is more preferred by one gender over the other.

# Results
```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-pretest
#| tbl-cap: "The results of the pretest."

htmltools::includeHTML(file.path(here::here(),"other/tables/table2.html"))
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-posttest
#| tbl-cap: "The results of the posttest."

htmltools::includeHTML(file.path(here::here(),"other/tables/table3.html"))
```

The results from pretest(@tbl-pretest) are pretty such the same for each group which aligns with the fact that this study was given before the experiment began. When looking at the post test results(@fig-boxplot), the control group is lagging behind the other groups, especially in word processing, spreadsheets and databases.

In word processing post test(@fig-boxplot), the group that perform the best were the social network and social gamifcation and roughly having the same performance. And then followed by educational game and gamification plugin. So it seems for word processing that groups with social elements are performing better than the group with game elements.

In spreadsheets post test(@fig-boxplot), the group that perform the best was the social gamification, while the other three groups perform about the same and was only little bit behind social gamification.

In presentations post test(@fig-boxplot), the group that perform the best was the educational game and not far behind was social network and gamification plugin. Surprisingly, social gamification was behind those three groups, as it would seem that the combination of social and game elements would make better than social network and gamification plugin.

In database post test(@fig-boxplot), the educational game performed as worse the control group. And similarly to the presentation posttest the social gamification group does worse than the social network and gamification plugin. And gamification plugin does slightly better than social network.

For the final examination(@fig-final-examination), interestingly enough, the group with the best results was the control group. Followed closely, by the social network group and then Educational game group. So the groups that performed the worse was gamfication plugin group and social gamfication group.

Our model results are summarized in @tbl-modelresults. Social gamification group grades are almost comparable to the control group as on average they tend to be lower by 1-2%. While, a student in the educational game group, their grades on average are 6-7% lower than the control group. And then, a student in social network group, their grades on average 11-12% lower than the control group. Lastly, a student in the gamification plugin group, their grades on average are 15-16% lower than the control group.

Males overall on average did better than females by 2-3%. Also, males in the gamification plugin group on average did better that females by 1-2%.  Interestingly, males do 2-4% worse than females in educational game, social gamification and social network group.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-boxplot
#| fig-cap: "Boxplot showing the results of the post test for control and experiment groups "
knitr::include_graphics(file.path(here::here(),"other/figures/figure4.png"))

```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-final-examination
#| fig-cap: "Boxplot showing the results of the final examination for each group that particapte in this experiment"
knitr::include_graphics(file.path(here::here(),"other/figures/figure5.png"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Final Exam marks based off of the experiment group and gender"
#| warning: false
library(rstanarm)

final_exam_model <-
  readRDS(file = here::here("models/final_exam.rds"))

modelsummary::modelsummary(
  list(
    "Final Exam Marks" = final_exam_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## Is Gamification effective? And is it better for male or females?

First, let us take a look at post test(@tbl-posttest) table again, recall that social gamification and social network does the best at word processing. Furthermore, social gamification does the best at spreadsheets as well. In addition, educational game does the best in presentations and gamification plugin does the best in databases. Notice that for post test, the best groups in each assessment seems to be related to groups that have game elements. Another thing to highlight is that social network group when comes to post tests tend to always be in top three groups. And social gamifcation is the best in 2/4 post tests. So it so seems that groups with social elements also perform well.

Gamification is not effective as we anticipated as seen in the final exam marks where the control group did the best. But, when comes to post tests as we note earlier groups with game elements were doing the best. And even for final examination social gamification was doing almost as good as the control group. From this study, we can conclude that gamification may not be good as we hope but it has potential to be better than traditional education. While, from our study we found males perform slightly better than females, the difference is pretty small(2-4%), which is good since both males and females can benefit equally from gamification. Furthermore, in order to improve gamification we will need to look at two things the design and student's feedback.

## Design of Gamification

Like I mentioned before gamification has to potential to change the traditional educational system, but in order to do that need to design it in a way that truly benefits the students. A leading expert in gamification and behavioral drive, Yu-kai Chou, has a create framework for gamification called the Octalysis Framework(reference to the figure in the appendix).Looking at the framework there are eight core drives: Epic Meaning, Accomplishment, Empowerment, Ownership, Social Influence, Scarcity, Unpredictability, and Avoidance. For even more in depth explanation, please visit Yu-kai website:https://yukaichou.com/gamification-examples/octalysis-complete-gamification-framework/.

Let us see how this framework applies to our study, for this example we will using social gamification group, since it one of the better performing groups with game elements. Epic Meaning category it is might but some students as they believe if they do well in this course, it could lead to some future goal like their dream job. Accomplishment is achieved by the meaningful challenges provided in each module and achievements. Empowerment is achieved by peer reviews and feedback from the teacher as well. Ownership is achieved as they can customize their profiles and status. Social Influence is achieved by being able to add friends and being able to features similar to Twitter(known as X now). This study has 5 out of 8 core drives mention from the framework.

Another example that we want to apply this framework is on a popular gamification language learning app, called Duolingo. Duolingo is a free app that has millions of user ranging from public students in third world countries to even Hollywood stars. Yu-kai Chou website provides an analysis for Duolingo, which has 6 out of 8 elements of the framework and here I am going to highlight some of the elements. Firstly, Epic Meaning is to learn a language and the app even asks the user when creating an account what is their purpose for learning new language. Secondly, Ownership, users have a currency call Lingots which allows players to buy Power-Ups(ways to make the player stronger) such as streak freeze and to buy clothes for their player character. Lastly, it has Scarcity and Avoidance as player have a limited amount of lives and if they lose then they can no longer play for a certain of time.

## Limations and Weaknessess
Limitations of this dataset is that doesn't tell us how student's satisfaction, retention, motivation and engagement. This is would useful information since it will paint us better picture on the effectiveness of gamification. In addition, if students don't really like then it would be indicator whether gamification is not working or that something missing that could improve it. According to study about academic boredom (@article), some aspects of it comes from lack of motivation and learning methods, so if gamification can solve this would a huge win for students and for society at a whole. Furthermore, if students really enjoy gamification then this means this a potential valid learning method, which could convince others to fund and study further.

Weakness of this dataset is that age column take the average age, which is 19 years in this case and gives every student that age. If the dataset gave us every student's age or even age range this could useful in determining which age group was benefiting more from gamification. It could that told us younger students better with gamification, but old students preform better with the traditional learning system.

## Future Studies

Like I mentioned before, this paper shows that gamification has potential to surpass traditional educational system. In future studies, more attention should be put into designing the gamification aspect. The Octalysis framework should be used and the design should it is best to meet all 8 core drives. Furthermore, Yu-kai is open to collaboration and consulting so that could make the design even more robust as there is more than just this framework that he talks on his website. In addition, there should focus on social elements as well since that experiement did well in post tests. Lastly, this design is probably going to need multiple iterations and constant feedback from the students to further improve it and to allow it eventually surpass the traditional education system.
\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]

plot(final_exam_model, "trace")

plot(final_exam_model, "rhat")
```

\newpage

# References

---
nocite: '@*'
---